---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
html_document:
keep_md: true
date: "Thursday, June 11, 2015"
---


## Loading and preprocessing the data


```{r ReadData}
activity <- read.csv("../activity.csv")

## Get the steps sum for each day
stepsEachDay<-tapply(activity$steps, format(as.Date(activity$date),'%Y-%m-%d'), sum) 

```

## Plot the histogram for the average daily steps
```{r histogram}
## Histogram for the average daily steps
hist(stepsEachDay, xlab="Average daily steps")

```



## What is mean total number of steps taken per day?
```{r MeanAndMedian}

##Get the mean for average daily steps
mean(stepsEachDay,na.rm=TRUE)

## Get the median for average daily steps
median(stepsEachDay,na.rm=TRUE)

```


## What is the average daily activity pattern?
``` {r dailypattern}
dfPlotInterval <- aggregate(steps ~ interval, activity, mean, na.action = na.omit)
plot(dfPlotInterval$interval, dfPlotInterval$steps, type='l', xlab="Interval", ylab="Number of steps",xaxt='n')
axis(side = 1, at=c(seq(from=0,to=2355,by=100)),labels = TRUE)
```

## Average maximum number of steps in the 5 minute interval
```{r averageMax}
maxRow <- dfPlotInterval[which.max(dfPlotInterval[,2]),]

```

The time series plot reveals the row `r maxRow$interval ` on average has maximum number of steps


## Imputing missing values
```{r}
nasfound <- nrow(activity[!complete.cases(activity),])

```
The total number of missing values or NAs `r nasfound`

```{r impute}
## The strategy devised is to make use of the data.frame that calculates the average of the 5 minute interval

# Get complete cases from activity data frame
tobeupdatedactivity <- activity[complete.cases(activity),]

##Get the mean for the 5 minute interval averaged across all days
meanSteps <- aggregate(steps ~ interval, activity, mean, na.action = na.omit)

## impute the missing values
impute.v <- function(x,p) {
  dat <- subset(activity, activity$interval==p & is.na(activity$steps))
  rowI <- meanSteps[x,]
  dat <- replace(dat, 1,rowI$steps)
  return(dat)
}

## Loop through and replace 

for(i in 1:288) {
  getRow <- meanSteps[i,]
  replaceddata <- impute.v(i,getRow$interval)
  tobeupdatedactivity <- rbind(replaceddata,tobeupdatedactivity)
}

## Histogram/mean/median for original data frame with imputed data
stepsImputedData<-tapply(tobeupdatedactivity$steps, format(as.Date(tobeupdatedactivity$date),'%Y-%m-%d'), sum) 
```

```{r histogramImputeData}

hist(stepsImputedData, xlab="Steps with Imputed data")

``` 

```{r meanImputedData}
mean(stepsImputedData)
median(stepsImputedData)

```


## Are there differences in activity patterns between weekdays and weekends?
```{r}

library(lubridate)

weekdayPlot <- aggregate(steps ~ interval, data=subset(tobeupdatedactivity,wday(tobeupdatedactivity$date) > 1 & wday(tobeupdatedactivity$date) < 7 ), mean)

weekendPlot <- aggregate(steps ~ interval, data=subset(tobeupdatedactivity,wday(tobeupdatedactivity$date) == 1 | wday(tobeupdatedactivity$date) == 7 ), mean)

weekdayPlot$DaysofWeek <- c("weekday")
weekendPlot$DaysofWeek <- c("weekend")

daysofweekDF <- rbind(weekdayPlot,weekendPlot)

```

```{r plot weekday/weekend }
library(ggplot2)
p <- ggplot(daysofweekDF, aes(interval, steps)) + geom_line()
p + facet_grid(DaysofWeek ~ .)

```


